Block bootstrap consistency\\under weak assumptions

### 2015-12-31 Paper.tex
[#6] Add further explanation for (old) Equation 18
This commit adds some further explanation for one of the equations
that a referee thought was not obvious.

For both results, we only present a proof for the stationary
bootstrap. The moving blocks and circular block bootstrap follow the
same general argument but are simpler. First we will define some
notation that will be used repeatedly.

Define the \sigmafield
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
$X_{nt}$. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays without modification to construct almost sure bounds on the
conditional moments $\Em$.

Also define
\begin{equation}
  \label{eq:10}
  I_n(\tau, m) = \begin{cases}
    \{\tau + 1,\dots, \tau + m\} &
    \textif\ 0 \leq \tau \leq n - m \text{\ and\ } 1 \leq m \\
    \{1,\dots, m - n + \tau\} \cup \{\tau + 1,\dots, n\} &
    \textif\ n - m < \tau \leq n \text{\ and\ } 1 \leq m\\
    \emptyset & \textif\  m \leq 0,
  \end{cases}
\end{equation}
so each $I_n(\tau, m)$ defines a potential block of length $m$ of the
original observations that could be chosen by the bootstrap.%
\footnote{The index sets $I_n(\tau, m)$ are designed to ``wrap
  around'' and use the first observations when $\tau + m > n$,
  matching the defining aspect of the stationary and circular block
  bootstraps.} %
By convention, summations over empty index sets will be considered
equal to zero. Also let
\newcommand{\Zb}[1][j]{{Z^*_{n{#1}}}}
\begin{align}\label{eq:11}
  Z_n(\tau, m) &= \tfrac{1}{\sqrt{n}} \sum_{t\in I_n(\tau, m)} (X_{nt} - \bar X_n) \\
  Z_n^*(\tau,m)&= \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau,m)} (X_{nt}^{*} - \bar X_n) \label{eq:12}
\intertext{and}
  \Zb          &= Z_n^*(K_{n,j-1}, M_{nj}),
\intertext{and define the corresponding demeaned terms}
  Z_n'(\tau, m) &= \tfrac{1}{\sqrt{n}} \sum_{t\in I_n(\tau, m)} (X_{nt} - \mu_{nt})
  \label{eq:13}\\
  Z_n^{\prime*}(\tau,m)&= \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau,m)} (X_{nt}^{*} - \mu_{nt}^*)
  \label{eq:14}
\intertext{and}
  Z_{nj}^{\prime *} &= Z_n^{\prime*}(K_{n,j-1}, M_{nj}), \label{eq:15}
\end{align}
where $\mu_{nt}^*$ is the expected value of the observation in the
original dataset corresponding to the $t$th observation in the
bootstrapped dataset. Further, define the filtration
\begin{equation}
\Gs_{nj} = \sigma(Z^*_{n1},\dots, Z^*_{nj},~X_{n1},\dots,X_{nn},~\Ms)
\end{equation}
so that $\{Z_{nj}^* / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array.

By construction,
\begin{equation}\label{eq:16}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^{*} - \bar X_n)
  = \sum_{j=1}^{J_n} \Zb
\end{equation}
and
\begin{equation}\label{eq:17}
  \Em^{*}g(\Zb,\dots,\Zb[k]) = \frac{1}{n^{k-j+1}}
  \sum_{\tau_1=0}^{n-1} \cdots \sum_{\tau_{k-j+1}=0}^{n-1}
  g(Z_n(\tau_1, M_{nj}),\dots,Z_n(\tau_{k-j+1}, M_{nk}))
\end{equation}
almost surely for any function $g$ and any $j \leq k$. Equation~\eqref{eq:17}
conditions on the lengths of each block, but averages over their
starting points.

>>>>

For both results, we only present a proof for the stationary
bootstrap. The moving blocks and circular block bootstrap follow the
same general argument but are simpler. We will define some additional
notation here before presenting the proofs.

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
$X_{nt}$. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays without modification to construct almost sure bounds on the
conditional moments $\Em$.

Also define
\begin{equation}
  \label{eq:10}
  I_n(\tau, m) = \begin{cases}
    \{\tau + 1,\dots, \tau + m\} &
    \textif\ 0 \leq \tau \leq n - m \text{\ and\ } 1 \leq m \\
    \{1,\dots, m - n + \tau\} \cup \{\tau + 1,\dots, n\} &
    \textif\ n - m < \tau \leq n \text{\ and\ } 1 \leq m\\
    \emptyset & \textif\  m \leq 0,
  \end{cases}
\end{equation}
so each $I_n(\tau, m)$ defines a potential block of length $m$ of the
original observations that could be chosen by the bootstrap.%
\footnote{The index sets $I_n(\tau, m)$ are designed to ``wrap
  around'' and use the first observations when $\tau + m > n$,
  matching the defining aspect of the stationary and circular block
  bootstraps.} %
By convention, summations over empty index sets will be considered
equal to zero. Note that the $I_n$ satisfy
\begin{align}\label{eq:4}
   \begin{split}
   I_n(K_{n0}, M_{n1}) &= \{1,\dots,K_{n1}\} \\
   I_n(K_{n1}, M_{n2}) &= \{K_{n1}+1,\dots,K_{n2}\} \\
   \vdots \\
   I_n(K_{n,J_n-1}, M_{nJ_{n}}) &= \{K_{n,J_{n-1}}+1,\dots,n \},
   \end{split}
\end{align}
so $I_n(K_{n0}, M_{n1}),\dots,I_n(K_{n,J_n-1}, M_{nJ_{n}})$ exactly
partition the set $\{1,\dots,n\}$ into consecutive blocks with
lengths determined by the bootstrap.

Let
\newcommand{\Zb}[1][j]{{Z^*_{n{#1}}}}
\begin{align}\label{eq:11}
  Z_n(\tau, m) &= \tfrac{1}{\sqrt{n}} \sum_{t\in I_n(\tau, m)} (X_{nt} - \bar X_n) \\
  Z_n^*(\tau,m)&= \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau,m)} (X_{nt}^{*} - \bar X_n) \label{eq:12}
\intertext{and}
  \Zb          &= Z_n^*(K_{n,j-1}, M_{nj}) = \tfrac{1}{\sqrt{n}} \sum_{t=K_{n,j-1}+1}^{K_{n,j}} (X_{nt}^* - \bar X_n) \label{eq:1}
\intertext{and define the corresponding demeaned terms}
  Z_n'(\tau, m) &= \tfrac{1}{\sqrt{n}} \sum_{t\in I_n(\tau, m)} (X_{nt} - \mu_{nt})
  \label{eq:13}\\
  Z_n^{\prime*}(\tau,m)&= \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau,m)} (X_{nt}^{*} - \mu_{nt}^*)
  \label{eq:14}
\intertext{and}
  Z_{nj}^{\prime *} &= Z_n^{\prime*}(K_{n,j-1}, M_{nj}), \label{eq:15}
\end{align}
where $\mu_{nt}^*$ is the expected value of the observation in the
original dataset corresponding to the $t$th observation in the
bootstrapped dataset. Further, define the filtration
\begin{equation}
\Gs_{nj} = \sigma(Z^*_{n1},\dots, Z^*_{nj},~X_{n1},\dots,X_{nn},~\Ms)
\end{equation}
so that $\{Z_{nj}^* / \sigma_n^*, \Gs_{nj}\}$ is a martingale difference
array.

By construction, (see Equations~\eqref{eq:4} and~\eqref{eq:1})
\begin{equation}\label{eq:16}
  \tfrac{1}{\sqrt{n}} \sum_{t=1}^n (X_{nt}^{*} - \bar X_n)
  = \sum_{j=1}^{J_n} \Zb
\end{equation}
and
\begin{equation}\label{eq:17}
  \Em^{*}g(\Zb,\dots,\Zb[k]) = \frac{1}{n^{k-j+1}}
  \sum_{\tau_1=0}^{n-1} \cdots \sum_{\tau_{k-j+1}=0}^{n-1}
  g(Z_n(\tau_1, M_{nj}),\dots,Z_n(\tau_{k-j+1}, M_{nk}))
\end{equation}
almost surely for any function $g$ and any $j \leq k$. Equation~\eqref{eq:17}
conditions on the lengths of each block, but averages over their
starting points.

### 2015-12-31 subs/ETr2/reply_2.tex

@@

@@

@@

>>>>

\section*{General summary of changes}
This version of the paper has the following changes from the
previous submission:
\input{summary}

\section*{Reply to specific points}

@@

  No, that was a mistake and it has been fixed. Thank you for pointing
  it out.

@@

  Thanks!

@@

  Each $Z_{nj}^*$ equals $(1/\sqrt{n}) \sum_{t=K_{n,j-1}+1}^{K_{nj}}
  (X_{nt}^* - \bar X_n)$ since $I_n(K_{n,j-1}, M_{nj}) =
  \{K_{n,j-1}+1,\dots,K_{nj}\}$. So the summations are taken over
  sequential and nonoverlapping blocks of the index $\{1,\dots,n\}$,
  starting with $\{1,\dots,M_{n1}\}$ and ending with
  $\{K_{n,J_n-1}+1,\dots,n\}$.  I've added some more details after the
  definition of $I_n(\tau,m)$ and spelled out the
  definition of $Z_{nj}^*$ a bit more to try to make
  this point more clearly in the paper. Be aware that the exact
  Equation number has moved.
